### 2014-11-21 Studies/2013_PG_County/Report/PG_Co_Fire.tex
PG Co: update report, plotting script, FDS output data and input file

\begin{table}[!h]
\centering
\captionof{table}[Abridged approximate fire event timeline]{Abridged Prince George's Fire/EMS approximate fire event timeline~\cite{PGCounty2013}.}\label{tab:fire_info}
\begin{tabular}{cl}
\toprule[1.5pt]
Incident Time              &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}               &                                        \\
\midrule
21:08:26   & First 911 call is made. \\[.25cm]
\multirow{3}{*}{21:11:03}  &   \multirow{3}{*}{\parbox{10cm} {Dispatch for a ``box'' alarm assignment- four engine companies, two truck companies, one rescue squad, and a commanding officer.}}    \\
                           &         \\
                           &         \\[.25cm]
21:12:55  &  First unit (Engine 807B) arrived on-scene.  \\[.25cm]
21:13:44  & Volunteer Chief 809A establishes incident command (IC).   \\[.25cm]
21:13:56  & Truck 809 arrived on-scene. \\[.25cm]
21:14:00  & Front door opened. \\[.25cm]
21:14:42  & Truck 809 enters structure. \\[.25cm]
21:15:47  & Basement door opened and water applied to fire. Front door closed over. \\[.25cm]
21:15:51  & Lower pane of front egress window broke. \\[.25cm]
21:16:01  & Top pane of front egress window broke. \\[.25cm]
21:16:16  & Truck 809 officer bailout through window. \\[.25cm]
21:16:44  & Front door opened. \\[.25cm]
21:17:15  & Truck 809 forcible entry firefighter rescued. \\[.25cm]
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\section{Fire}
\label{fire}
To estimate the type of fuel, fire size, and structure ventilation, on-scene videos and post-incident reports were used. Based on these sources of information, the FDS model includes two source fire locations in the basement: one fire in the kitchen area and one in the adjecent bathoom (Fig.~\ref{fig:basement_fire}). At the time of the fire, the structure was vacant so the the dominant fuel was wood: exposed ceiling joists, subfloor, cabinets, and stairs.

@@

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The first assumption is that the fires start at the beginning of the simulation and increase to a value of 4.5~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations (1 source for the kitchen, 1 source for the bathroom). The following lines define the source fires in the FDS input file:

\begin{lstlisting}
&SURF ID='burner1', HRRPUA = 1000, COLOR='BURNT ORANGE', RAMP_Q='BURN' /
&SURF ID='burner2', HRRPUA = 2000, COLOR='BURNT ORANGE', RAMP_Q='BURN' /

&RAMP ID='BURN', T = 0,   F = 0.  /
&RAMP ID='BURN', T = 10,  F = 1.0 /
&RAMP ID='BURN', T = 300, F = 1.0 /

&OBST XB=0.5,1.0,5.5,6.5,0.0,0.4, SURF_IDS='burner1','INERT','INERT' /  Burner1
&OBST XB=2.3,4.3,5.5,6.5,0.0,0.4, SURF_IDS='burner2','INERT','INERT' /  Burner2
\end{lstlisting}
Note that the coordinates here are unique for the vents are specific to the input files used for these simulations. The key point is the 2-D planar area.

In reality, the quantities associated with these input parameters are not fixed values; rather, a model input parameter can be thought of as a point estimate from a distribution of possible input parameters with some associated amount of uncertainty. Any change in an input parameter (such as the HRR) for a given scenario would result in a change in the output quantity (such as the hot gas layer (HGL) temperature). For example, according to the McCaffrey, Quintiere, and Harkleroad~\cite{SFPE:Walton} empirical correlation, the HGL temperature in a well-ventilated compartment fire is proportional to the HRR raised to the two-thirds power. Following this relationship, a 7.5~\% increase in the HRR would result in a 5~\% increase in the HGL temperature~\cite{NUREG_1824_Sup_1}. More detailed discussion on the propagation of parameter uncertainty in fire models is available in a validation study that was sponsored by the U.S. Nuclear Regulatory Commission~\cite{NUREG_1824_Sup_1}. The following subsections describe the inputs that were used to develop the simulation, including any approximations or assumptions that were made.

@@

\section{Numerical Mesh}
\label{mesh}
For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm (3.9~in), the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 7 for the 0.5~MW bathroom fire and 17 for the 4.0~MW basement fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

@@

Figure~\ref{fig:geom_grid} shows the front and rear of the structure and surrounding domain rendered in Smokeview with the 10~cm and 20~cm meshes. Note the finer mesh (smaller cells) surrounding the structure and the coarser mesh on each side. The domain was divided into 12 equally sized 10~cm meshes that surround the structure (each containing 150,000 grid cells) plus 4 equally sized 20~cm meshes to the left and right of the stucture (each containing 32,000 grid cells) and one 20~cm mesh in the rear of the structure (containing 56,000 grid cells) for a total of 17 meshes. Multiple meshes allowed the simulation to be processed in parallel, which reduced the amount of required calculation time to approximately 18~hours for the no wind case and 3~days for the 20~mph wind case. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. The boundary between the 12 10~cm meshes and 5 20~cm meshes is indicated by the {\bf bold} lines. Table~\ref{tab:mod_param} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

@@

\begin{table}
\centering
\captionof{table}{Relevant fire model input parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter                                                  & Description                                   & Discussion                            \\
\midrule
Simulation Time                                            &  5~min                                        & --                                    \\ [.25cm]
\multirow{2}{*}{Grid Cell Size}                            &  10~cm around house                           & \multirow{2}{*}{Section~\ref{mesh}}   \\ 
                                                           &  20~cm for wind                               &                                       \\ [.25cm]
Ambient Temperature$^*$                                    &  20~$^{\circ}$C (68~$^{\circ}$F)              & --                                    \\ [.1cm]
\multirow{4}{*}{Reaction: Wood~\cite{SFPE:Tewarson}}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     & \multirow{4}{*}{Section~\ref{fire}}   \\
                                                           &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                       \\
                                                           &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                       \\
                                                           &  Heat of Combustion: 16,400~kJ/kg             &                                       \\ [.25cm]
\multirow{2}{*}{Peak HRR}                                  &  Basement Fire: 6.75~MW                       & \multirow{2}{*}{Section~\ref{fire}}   \\ 
                                                           &  Bathroom Fire: 1.8~MW                        &                                       \\ [.25cm]                     
\multirow{3}{*}{Material: Wood~\cite{Incropera:1} }        &  $k$: 0.12~\si{W/(m.K)})                      & \multirow{3}{*}{Section~\ref{matl}}   \\
                                                           &  $\rho$: 510~kg/m$^3$                         &                                       \\
                                                           &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                       \\ [.25cm]
\multirow{3}{*}{Material: Gypsum Board~\cite{WAKILI2007}}  &  $k$: 0.28~\si{W/(m.K)}                       &  \multirow{3}{*}{Section~\ref{matl}}  \\ 
                                                           &  $\rho$: 810~kg/m$^3$                         &                                       \\ 
                                                           &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                       \\
\bottomrule[1.25pt]
\end{tabular}\par
\footnotesize
$^{*}$Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 10~$^{\circ}$C~\cite{PGCounty2013}.
\normalsize
\end{table}

>>>>

\begin{table}[!h]
\centering
\captionof{table}[Abridged approximate fire event timeline]{Abridged Prince George's Fire/EMS approximate fire event timeline~\cite{PGCounty2013}.}\label{tab:fire_info}
\begin{tabular}{cl}
\toprule[1.5pt]
Incident Time              &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}               &                                        \\
\midrule
21:08:26   & First 911 call is made. \\[.25cm]
\multirow{3}{*}{21:11:03}  &   \multirow{3}{*}{\parbox{10cm} {Dispatch for a ``box'' alarm assignment- four engine companies, two truck companies, one rescue squad, and a commanding officer.}}    \\
                           &         \\
                           &         \\[.25cm]
21:12:55  &  First unit (Engine 807B) arrived on-scene.  \\[.25cm]
21:13:44  & Volunteer Chief 809A establishes incident command (IC).   \\[.25cm]
21:13:56  & Truck 809 arrived on-scene. \\[.25cm]
21:14:00  & Front door opened. \\[.25cm]
21:14:42  & Truck 809 enters structure. \\[.25cm]
21:15:47  & Basement door opened and water applied to fire. Front door blown closed. \\[.25cm]
21:15:51  & Lower pane of front egress window broke. \\[.25cm]
21:16:01  & Top pane of front egress window broke. \\[.25cm]
21:16:16  & Truck 809 officer bailout through window. \\[.25cm]
21:16:44  & Front door opened. \\[.25cm]
21:17:15  & Truck 809 forcible entry firefighter rescued. \\[.25cm]
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\section{Fire}
\label{fire}
To estimate the type of fuel, fire size, and structure ventilation, on-scene videos and post-incident reports were used. Based on these sources of information, the FDS model includes two source fire locations in the basement: one fire in the kitchen area and one in the adjecent bathoom (Fig.~\ref{fig:basement_fire}). At the time of the fire, the structure was vacant so the the dominant fuel was wood: wood paneling on walls, exposed ceiling joists, subfloor, cabinets, and stairs.

@@

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The first assumption is that the fires start at the beginning of the simulation and increase to a value of 8.55~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. Since the actual ignition time is unknown, the fire was ramped quickly to reflect the conditions determined from reports and post-incident images. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations in the basement (1 source for the kitchen, 1 source for the bathroom).  At 207~s, using a simple decay model, the fire size decreases lineraly to 25~\% of its peak value in 60~s. This is to account for the impact of firefighters putting water on the basement fire (Table~\ref{tab:fire_info}). The following lines define the source fires in the FDS input file:

\begin{lstlisting}
&SURF ID='burner1', HRRPUA = 2000, COLOR='BURNT ORANGE', RAMP_Q='BURN' /
&SURF ID='burner2', HRRPUA = 4500, COLOR='BURNT ORANGE', RAMP_Q='BURN' /

&RAMP ID='BURN', T = 0,   F = 0.  /
&RAMP ID='BURN', T = 10,  F = 1.0 /
&RAMP ID='BURN', T = 207, F = 1.0 /
&RAMP ID='BURN', T = 267, F = 0.25 /
&RAMP ID='BURN', T = 300, F = 0.25 /

&OBST XB=0.6,1.2,5,6.5,0.0,0.4, SURF_IDS='burner1','INERT','INERT' /  Burner1
&OBST XB=2.5,4,5.0,6.0,0.0,0.4, SURF_IDS='burner2','INERT','INERT' /  Burner2
\end{lstlisting}
Note that the coordinates here are unique for the vents are specific to the input files used for these simulations. The key point is the 2-D planar area.

Figure~\ref{fig:hrr} shows the overall prescribed design fire for this scenario, which is the sum of the HRRs of the three source fires. In this figure, the text labels the prescribed steady peak HRR and the decay phase associated with firefighter water application. Therefore, the estimated maximum specified fire size for this structure was approximately 8.25~MW. Note that the prescribed HRR can be different than the simulated HRR calculated by FDS based on the amount of available fuel, oxygen, and ventilation. A comparison of the prescribed HRR and the simulated HRR is discussed in Section~\ref{HRR}.

\begin{figure}[!ht]
\includegraphics[width=0.75\textwidth]{../Figures/Fire_HRR}
\caption[Prescribed HRR vs. time for the simulation.]
{Prescribed HRR vs. time for the simulation. The text labels indicate the prescribed steady peak HRR and the decay phase associated with firefighter water application.}
\label{fig:hrr}
\end{figure}

\clearapge

@@

\section{Numerical Mesh}
\label{mesh}
For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm (3.9~in), the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 12 for the 1.8~MW bathroom fire and 20 for the 6.75~MW basement fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

@@

Figure~\ref{fig:geom_grid} shows the front and rear of the structure and surrounding domain rendered in Smokeview with the 10~cm and 20~cm meshes. Note the finer mesh (smaller cells) surrounding the structure and the coarser mesh on each side. The domain was divided into 12 equally sized 10~cm meshes that surround the structure (each containing 150,000 grid cells) plus 4 equally sized 20~cm meshes to the left and right of the stucture (each containing 32,000 grid cells) and one 20~cm mesh in the rear of the structure (containing 56,000 grid cells) for a total of 17 meshes. Multiple meshes allowed the simulation to be processed in parallel, which reduced the amount of required calculation time to approximately 18~hours for the no wind case and 3~days for the 20~mph wind case. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. The boundary between the twelve 10~cm meshes and five 20~cm meshes is indicated by the {\bf bold} lines. Table~\ref{tab:mod_param} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

@@

\section{Summary of Model Input Parameters}
\label{sec:summary_of_model_inputs}

Table~\ref{tab:mod_param} shows a summary of the model input parameters for the simulation that was conducted as part of this study. In reality, the quantities associated with model input parameters are not fixed values; rather, a model input parameter can be thought of as a point estimate from a distribution of possible input parameters with some associated amount of uncertainty. Any change in an input parameter (such as the HRR) for a given scenario results in a change in the output quantity (such as the hot gas layer (HGL) temperature).

For example, according to the McCaffrey, Quintiere, and Harkleroad~\cite{SFPE:Walton} empirical correlation, the HGL temperature in a well-ventilated compartment fire is proportional to the HRR raised to the two-thirds power. Following this relationship, a 7.5~\% increase in the HRR would result in a 5~\% increase in the HGL temperature~\cite{NUREG_1824_Sup_1}. More detailed discussion on the propagation of parameter uncertainty in fire models is available in a validation study that was sponsored by the U.S. Nuclear Regulatory Commission~\cite{NUREG_1824_Sup_1}.

\begin{table}[!ht]
\centering
\captionof{table}{Relevant fire model input parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter                                                  & Description                                   & Discussion                            \\
\midrule
Simulation Time                                            &  5~min                                        & --                                    \\ [.25cm]
\multirow{2}{*}{Grid Cell Size}                            &  10~cm around house                           & \multirow{2}{*}{Section~\ref{mesh}}   \\ 
                                                           &  20~cm for wind                               &                                       \\ [.25cm]
Ambient Temperature$^*$                                    &  20~$^{\circ}$C (68~$^{\circ}$F)              & --                                    \\ [.1cm]
\multirow{4}{*}{Reaction: Wood~\cite{SFPE:Tewarson}}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     & \multirow{4}{*}{Section~\ref{fire}}   \\
                                                           &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                       \\
                                                           &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                       \\
                                                           &  Heat of Combustion: 16,400~kJ/kg             &                                       \\ [.25cm]
\multirow{2}{*}{Peak HRR}                                  &  Basement Fire: 6.75~MW                       & \multirow{2}{*}{Section~\ref{fire}}   \\ 
                                                           &  Bathroom Fire: 1.8~MW                        &                                       \\ [.25cm]                     
\multirow{3}{*}{Material: Wood~\cite{Incropera:1} }        &  $k$: 0.12~\si{W/(m.K)})                      & \multirow{3}{*}{Section~\ref{matl}}   \\
                                                           &  $\rho$: 510~kg/m$^3$                         &                                       \\
                                                           &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                       \\ [.25cm]
\multirow{3}{*}{Material: Gypsum Board~\cite{WAKILI2007}}  &  $k$: 0.28~\si{W/(m.K)}                       & \multirow{3}{*}{Section~\ref{matl}}   \\ 
                                                           &  $\rho$: 810~kg/m$^3$                         &                                       \\ 
                                                           &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                       \\
\bottomrule[1.25pt]
\end{tabular}\par
\footnotesize
$^{*}$Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 10~$^{\circ}$C~\cite{PGCounty2013}.
\normalsize
\end{table}
